---
title: "Figure 1 — Buffa hypoxia scores across cancers and lung tissues"
author: "OCRscore manuscript pipeline"
output: html_document
---

**Standalone notebook** — does **not** require `targets`, `paths.json`, or the
T-drive / Box expression trees.

| Panel | Input |
|-------|-------|
| **A** | [`data/fig1a_buffa_tcga.rds`](data/fig1a_buffa_tcga.rds) (tracked) |
| **B** | [`data/fig1b_buffa_lung.csv`](data/fig1b_buffa_lung.csv) — add when available |

From `manuscript/`:

```r
rmarkdown::render("figure_1.Rmd")
# or step through chunks in the IDE
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(colorspace)
library(patchwork)
source("R/plot_buffa_raincloud.R")
```

# Panel A — Buffa across cancers (KIRC excluded)

```{r panel-a}
fig1a_dat <- readRDS("data/fig1a_buffa_tcga.rds") %>%
  mutate(cancer = toupper(cancer)) %>%
  filter(cancer != "KIRC") %>%
  mutate(mean = mean(BUFFA_HYPOXIA_SCORE), .by = cancer) %>%
  mutate(
    group = forcats::fct_reorder(cancer, -mean),
    score = BUFFA_HYPOXIA_SCORE
  )

p_a <- plot_buffa_raincloud(fig1a_dat)
p_a
ggsave("figures/Fig1A_buffa-cancer-raincloud.png", p_a, height = 7, width = 5)
ggsave("figures/Fig1A_buffa-cancer-raincloud.svg", p_a, height = 7, width = 5)
```

# Panel B — Lung normal vs tumor

Needs columns `source` and `buffa.score`. Short codes `luad` / `lusc` are
treated as adjacent-normal; labels become “Adjacent Normal LUAD”, “Healthy
Lung”, “LUAD”, etc.

```{r panel-b}
# Prefer the slim lung table; fall back to the full lab export if present
fig1b_candidates <- c(
  "data/fig1b_buffa_lung.csv",
  "data/buffa_avgZscore.csv"
)
fig1b_path <- fig1b_candidates[file.exists(fig1b_candidates)][1]
have_fig1b <- !is.na(fig1b_path)

if (!have_fig1b) {
  message(
    "Skipping Panel B — missing data/fig1b_buffa_lung.csv ",
    "(or data/buffa_avgZscore.csv).\n",
    "Needed columns: source, buffa.score."
  )
  p_b <- NULL
} else {
  message("Panel B input: ", fig1b_path)
  fig1b_dat <- read.csv(fig1b_path) %>%
    format_lung_buffa()

  p_b <- plot_buffa_raincloud(fig1b_dat)
  print(p_b)
  ggsave("figures/Fig1B_buffa-lung-normal-tumor-raincloud.png", p_b,
         height = 5, width = 6)
  ggsave("figures/Fig1B_buffa-lung-normal-tumor-raincloud.svg", p_b,
         height = 5, width = 6)
}
```

# Combined A + B

```{r combined}
if (is.null(p_b)) {
  message("Skipping combined Figure 1 — Panel B not available.")
} else {
  p_ab <- p_a + p_b +
    plot_layout(widths = c(1, 1)) +
    plot_annotation(tag_levels = "A") &
    theme(
      text = element_text(family = "Arial", size = 8),
      axis.title = element_text(family = "Arial", size = 8),
      axis.text = element_text(family = "Arial", size = 8),
      plot.tag = element_text(family = "Arial", face = "bold", size = 11)
    )

  ggsave("figures/Fig1_buffa-rainclouds.png", p_ab, height = 5, width = 6.5)
  ggsave("figures/Fig1_buffa-rainclouds.svg", p_ab, height = 5, width = 6.5)
}
```
