---
title: "Figure 3 — OCR score, hypoxia, and HIF targets"
author: "OCRscore manuscript pipeline"
output: html_document
---

Panels for Figure 3 only. Needs a local `data_root` with `many-cancers/`,
`tcga-expression/`, and `pancan_pcawg_2020/` (see `../../paths.example.json`).

```r
# copy ../../paths.example.json → ../../paths.json and set data_root, then:
rmarkdown::render("figure_3.Rmd")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggtext)
library(tmesig)
source("R/setup_paths.R")
source("R/readAndFilter.R")
source("R/loadOCRscores.R")
source("R/loadManyScores.R")
source("R/formatForSmooth.R")
source("R/loadAndFormatTCGA.R")
source("R/loadAndFormatPCAWG.R")

paths <- setup_paths()
ocrscores <- loadOCRscores(paths$many_cancers)
plotdat_scores_for_lm <- formatForSmooth(loadManyScores(paths$many_cancers))
integrated_TCGA_PCAWG <- bind_rows(
  loadAndFormatPCAWG(paths$pcawg),
  loadAndFormatTCGA(paths$tcga)
)
```

# Panel A — OCR score vs Buffa hypoxia score (all cancers)

Linear fits are stratified by Buffa positivity (score > 0 vs ≤ 0).

```{r panel-a}
ocrscores %>%
  ggplot(aes(x = BUFFA_HYPOXIA_SCORE, y = AvgExpr)) +
  geom_point(aes(group = pos), alpha = 0.2, color = "grey") +
  theme_bw() +
  stat_smooth(aes(group = pos), method = "lm", color = "red", lwd = 1.5) +
  labs(x = "Buffa Hypoxia Score",
       y = "OCR score")
ggsave("figures/Fig3A_scatterplot_buffa-OCR_all_split-lm.png",
       height = 3, width = 5)
```

# Panel B — OCR, mitochondrial biogenesis, and mitophagy in LB vs HB cancers

Representative low-Buffa (LB) and high-Buffa (HB) cancer types.

```{r panel-b}
plotdat_scores_for_lm %>%
  filter(cancer %in% c("THCA", "PRAD", "LGG", "CESC", "LUSC", "HNSC")) %>%
  mutate(`Buffa Category` = fct_relevel(`Buffa Category`, "LB"),
         signature = ifelse(grepl("Electron", signature), "OCR score", signature)) %>%
  ggplot(aes(x = BUFFA_HYPOXIA_SCORE, y = score)) +
  geom_point(color = "grey", alpha = 0.2) +
  geom_smooth(aes(color = signature), method = "lm") +
  facet_wrap(~ `Buffa Category` + cancer) +
  labs(x = "Buffa Hypoxia Score",
       y = "Average Log<sub>2</sub> Expression") +
  scale_color_viridis_d(direction = -1, name = "Signature") +
  theme_bw() +
  theme(axis.title.y = element_markdown(),
        legend.position = "top")

ggsave("figures/Fig3B_hb-lb_3sigs_lm_facet.png",
       height = 5,
       width = 5)
```

# Panel C — HIF-related genes / MIR210 vs Buffa (TCGA + PCAWG)

*PDK1*, *PDK3*, and *NDUFA4L2* from TCGA; *MIR210* (*hsa-miR-210-3p*) from PCAWG.

```{r panel-c}
integrated_TCGA_PCAWG %>%
  ggplot(aes(x = buffa.score, y = log.counts)) +
  facet_wrap(vars(Hugo_Symbol), nrow = 1) +
  geom_point(color = "grey", alpha = 0.2) +
  stat_smooth(method = "lm", color = "red") +
  labs(x = "Buffa hypoxia score",
       y = expression(paste(Log[2], "(Expression + 1)", sep = ""))) +
  theme_bw()
ggsave("figures/Fig3C_scatterplot_buffa-MIR210-genes_TCGA-pcawg.png",
       height = 3, width = 5)
```
