---
title: "Figure 4 — Core vs supernumerary subunit OS enrichment"
author: "OCRscore manuscript pipeline"
output: html_document
---

Computes permutation-test *P* values for Figure 4 (core vs supernumerary ETC
subunits × OS outcome).

**Standalone** — tracked Excel only.

```r
rmarkdown::render("figure_4.Rmd")
```

Input: `../../exploratory/data/Subunit OS 11242021.xlsx`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(readxl)
source("R/tallyArms.R")
source("R/make_comparison_data.R")

os_file <- "../../exploratory/data/Subunit OS 11242021.xlsx"
F4_OS_data <- read_xlsx(os_file, na = "N/A")
original_data <- tallyArms(F4_OS_data)
comparison_data <- make_comparison_data(F4_OS_data)
```

# Observed counts

```{r observed}
original_data
```

# Empirical one-sided *P* values (B = 1000 permutations)

*P* = proportion of permutations with count ≥ observed count.

```{r p-values}
pval <- function(obs, null_col) {
  sum(comparison_data[[null_col]] >= obs, na.rm = TRUE) / nrow(comparison_data)
}

results <- tibble::tribble(
  ~category,              ~observed,                              ~p_value,
  "Core - Positive",      original_data$Core.Positive,            pval(original_data$Core.Positive, "Core.Positive"),
  "Core - Negative",      original_data$Core.Negative,            pval(original_data$Core.Negative, "Core.Negative"),
  "Core - NA",            original_data$Core.NA,                  pval(original_data$Core.NA, "Core.NA"),
  "Supernumerary - Positive", original_data$Supernumerary.Positive, pval(original_data$Supernumerary.Positive, "Supernumerary.Positive"),
  "Supernumerary - Negative", original_data$Supernumerary.Negative, pval(original_data$Supernumerary.Negative, "Supernumerary.Negative"),
  "Supernumerary - NA",   original_data$Supernumerary.NA,         pval(original_data$Supernumerary.NA, "Supernumerary.NA")
)

print(results)
readr::write_csv(results, "figures/Fig4_core-supernumerary_permutation_pvalues.csv")
```

The structural Complex I/II schematic in the paper is assembled from this
curated table outside R.
