---
title: "lung buffa/mitoscore"
author: "C. Dravillas"
date: "2025-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tmesig)
library(purrr)

source("00-paths.R")
```

# Read in Data

```{r}
# n = 955
pt.key <- read.delim("../data/tho_lung_cancer.s_clinical_data.tsv") %>%
  select(Patient.ID, Sample.ID, Sequencing.Type) %>%
  separate(Sample.ID, into = c("Pt.id", "sample", "WES.ID"), sep = "_", fill = "right") %>%
  mutate(WES.ID = ifelse(Sequencing.Type == "WES", sample, WES.ID),
         sample = ifelse(Sequencing.Type == "WES", NA, sample),
         sample = gsub("-RNA", "", sample)) %>%
  select(Patient.ID, sample) %>%
  filter(!is.na(sample))

gene.path <- "TCC00601-BENE/RNAseq/gene_and_transcript_expression_results"

# n = 182 in overlapping sample set 
gene.files <- data.frame(files = list.files(file.path(paths$tdrive, gene.path), 
                                   full.names = TRUE)) %>%
  filter(grepl("genes", files)) %>%
  filter(!grepl("md5", files)) %>%
  mutate(sample = sub(".*/([^/]+)\\.genes\\.results$", "\\1", files),
         sample = sub("-RNA", "", sample)) %>%
  filter(sample %in% pt.key$sample)
```

There are 3 different sample formats 
SL-
AZSF
FT- 
Note: Not all sample present in clinical tsv are present on t-drive 

```{r}
expr <- gene.files %>%
  mutate(data = map2(files, sample, ~ read.delim(.x) %>% 
                        mutate(sample = .y))) %>%
  pull(data) %>%
  bind_rows()

head(expr)
```

expected count, TPM and FPKM available 
tmesig function uses counts 

need in format @param gene_matrix A data frame of expression counts where rows 
are genes, columns are samples, and the gene symbols are available in a column 
defined as "Gene".

Need to handle duplicate counts for SLC25A6, there is a version that is 0 for all,
will remove that one I guess? 

```{r}
expr.in  <- expr %>%
  dplyr::select(sample, gene_symbol, expected_count) %>%
  dplyr::rename(Gene = gene_symbol) %>%
  filter(Gene %in% inputGenes("Mitoscore") | Gene %in% inputGenes("Buffa"),
         !(Gene == "SLC25A6" & expected_count == 0)) %>%
  pivot_wider(names_from = sample, values_from = expected_count)
```

Calculate Mitoscore & Buffa

```{r}
mitoscore <- tmesig::calculateMitoscore(expr.in, mito.genes = inputGenes("Mitoscore"))
buffa <- tmesig::calculateBuffa(expr.in, buffa.genes = inputGenes("Buffa"))
```

Combine scores and save
```{r}
scores <- merge(mitoscore, buffa)

write.csv(scores, file = "../data/tho_lung_cancer-mitoBuffa.csv", row.names = FALSE)
```



