---
title: "Fig 3 Buffa and mitochondrial oxygen demand"
author: "Dan Spakowicz"
date: "11/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load required packages
library(tidyverse)
library(gt)
library(ggalt)
library(ggridges)
library(forcats)

# Source paths 
source("00-paths.R")
```

Code to create Figure 3 in __"Oxygen demand driven tumor hypoxia: a new perspective on the genesis and treatment of hypoxia"__

>__Figure 3. Schematic representation of relationship between hypoxia and mitochondrial oxygen demand.__ _Low mean Buffa score (LB)_ cancers show positive correlation between increasing hypoxia, mitoscore, mitochondrial biogenesis and mitophagy. _High mean Buffa score (HB)_ cancers show a more moderate correlation between hypoxia, mitoscore, and mitochondrial biogenesis with an inverse relationship with mitophagy. Lower slope mitoscore:Buffa score relationship in the more hypoxic tumors corresponds to elevated HIF-1 target gene expression carbonic anhydrase IX (CA9), and pyruvate dehydrogenase kinases 1 and 3 genes.


```{r}
x <- data.frame(files = list.files( file.path(paths$box, 
                                              "data", 
                                              "many-cancers"), 
                                    full.names = TRUE)) %>%
  filter(grepl("\\.txt", files)) %>%
  filter(grepl("MaxLimitsall", files))


mitophagy.files <- x %>%
  filter(grepl("GOBP_MITOPHAGY", files))

biogen.files <- 
  x %>%
  filter(grepl("BIOGEN", files))

etc.files <- 
  x %>%
  filter(grepl("ELECTRON", files))
```

```{r}
# Define a function to read in all the files and select columns of interest
readAndFilter <- function(table) {
  read.table(table) %>%
    select(BUFFA_HYPOXIA_SCORE, AvgExpr) %>%
    rownames_to_column("sampleID") %>%
    mutate(cancer = table) %>%
    mutate(cancer = gsub(pattern = ".*LogExpr_(.*)_AverageGene.*", 
                         replacement = "\\1",
                         x = cancer))
}

mitophagy <- lapply(X = mitophagy.files$files, readAndFilter) %>%
  bind_rows() %>%
  rename("Mitophagy" = "AvgExpr")

biogen <- lapply(X = biogen.files$files, readAndFilter) %>%
  bind_rows() %>%
  rename("Mitochondrial Biogenesis" = "AvgExpr")

etc <- lapply(X = etc.files$files, readAndFilter) %>%
  bind_rows() %>%
  rename("Electron Transport Chain" = "AvgExpr")

comb <- 
  mitophagy %>%
  full_join(biogen) %>%
  full_join(etc) %>%
  mutate(pos = if_else(BUFFA_HYPOXIA_SCORE > 0,
                      true = "positive",
                      false = "negative"))

dim(comb)
head(comb)
```
There's a concept of low and high mean buffa score cancers. I'll take that at face value for the assignment

```{r}
mean.buffa <- 
  comb %>%
  group_by(cancer) %>%
  summarize(mean = mean(BUFFA_HYPOXIA_SCORE)) %>%
  mutate(`Buffa Category` = if_else(mean > 0,
                             true = "HB",
                             false = "LB")) %>%
  mutate(cancer = fct_reorder(cancer, mean))

mean.buffa %>%
  gt()
```

```{r}
mean.buffa %>%
  ggplot(aes(x = mean, y = cancer)) +
  geom_segment(aes(y = cancer, yend = cancer, x = mean, xend = 1), size = .8, alpha = .8) +
  geom_point(aes(x = mean, y = cancer, fill = `Buffa Category`), shape = 21, size = 6, alpha = .9) +
  theme_bw() +
  labs(x = "Mean Buffa Score",
       y = "") +
  scale_fill_viridis_d()

ggsave("../figures/mean-buffa_lollipop.png",
       height = 4, width = 6)
```



```{r}
cmb <- 
  comb %>%
  full_join(mean.buffa) %>%
  mutate(cancer = fct_reorder(cancer, mean))

cmb %>%
  ggplot(aes(x = BUFFA_HYPOXIA_SCORE, y = cancer, fill = stat(x))) +
  geom_density_ridges_gradient(show.legend = FALSE) +
  theme_bw() +
  labs(x = "Buffa Hypoxia Score",
       y = "") +
  scale_fill_viridis_c()
ggsave("../figures/buffa_density-ridges.png",
       height = 4, width = 6)
```



```{r}
cmbl <- 
  cmb %>%
  select(-mean) %>%
  gather(key = "signature", value = "score", 
         -BUFFA_HYPOXIA_SCORE,
         -cancer,
         -pos,
         -`Buffa Category`,
         -sampleID)

head(cmbl)
```

```{r}
cmbl %>%
  ggplot(aes(x = BUFFA_HYPOXIA_SCORE, y = score)) +
  geom_point(color = "grey", alpha = 0.2) +
  geom_smooth(aes(color = signature)) +
  facet_wrap(~`Buffa Category`) +
  labs(x = "Buffa Hypoxia Score",
       y = "Mitochondrial Oxygen Demand") +
  theme_bw() +
  scale_color_viridis_d()
ggsave("../figures/Fig3_hb-lb_3sigs_loess.png",
       height = 4,
       width = 6)
```

