---
title: "mitoscore modeling"
author: "Caroline Wheeler"
date: '2022-11-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(MASS)

source("00-paths.R")
```

read in cna buffa table with 93 gene matches
```{r}
df <- readRDS("../data/cna_buffa.rds")
```

expr level ~ cna category

# need to split by cancer
```{r}
split_df <- split(df, f = df$cancer)  
```

function to model and save important output
```{r}
modelBuffa <- function(df){
  #df <- split_df$ucec
  
  gene <- df %>%
  dplyr::select(Hugo_Symbol) %>%
  distinct()
  
  geneNames <- gene$Hugo_Symbol

  output <- data.frame(matrix(ncol = 5, nrow = 0))
  #provide column names
  colnames(output) <- c('Value', 'Std. Error', 't value', 'p value', 'gene')
  
  for (g in geneNames) {
    temp <- df %>%
      dplyr::filter(Hugo_Symbol == g)
  
      tryCatch({
           # test the call in question
          m = polr(value ~ BUFFA_HYPOXIA_SCORE, data = temp, Hess = TRUE)
          ctable <- coef(summary(m))
          p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
          ctable <- cbind(ctable, "p value" = p)
  
          data <- data.frame(matrix(ncol = 1, nrow = 5))
          colnames(data) <- c("gene")
          data$gene <- as.character(g)
  
          ctable <- cbind(ctable, data)
         }, error = function(e) {
            # handle error event
           
            message(e)
            
            ctable <- data.frame(matrix(ncol = 5, nrow = 5))
            colnames(ctable) <- c('Value', 'Std. Error', 't value', 'p value', 'gene')
            ctable$gene <- as.character(g)
            ctable$Value <- NA
            ctable$`Std. Error` <- NA
            ctable$`p value`<- NA
            ctable$`t value` <- NA
         })
  
    output <- rbind(output, ctable)
    
  }
  
  filt_output <- output %>%
  filter(grepl("BUFFA", rownames(output)))

  filt_output$test <- ""
  filt_output$cancer <- df$cancer %>% unique()
  
  return(filt_output)
  
  #all_cancers <- rbind(all_cancers, filt_output)
}
```

Getting following error with COADREAD, GBM, KIRC, KIRP, PCPG, PRAD, THCA
Warning: glm.fit: algorithm did not converge
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Error in polr(value ~ BUFFA_HYPOXIA_SCORE, data = temp, Hess = TRUE) : 
  attempt to find suitable starting values failed

# using group by and do
```{r}
all_cancers <- data.frame(matrix(ncol = 7, nrow = 0))
colnames(all_cancers) <- c('Value', 'Std. Error', 't value', 'p value', 'gene', 'test', 'cancer')

for (data in split_df) {
  temp <- modelBuffa(data)
  all_cancers <- rbind(all_cancers, temp)
}
```

manually remove crazy outlier
```{r}
all_cancers <-subset(all_cancers, cancer!="thca" | gene!="COX5A")
```


```{r}
ggplot(all_cancers, aes(x=cancer , y=gene, fill=Value)) +
 geom_tile(color = "black") +
    scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000") +
  xlab("") +
  ylab("") 

ggsave("../figures/all_cancers_heatmap.png", dev = "png", width = 5, height = 10, dpi = 600)
```

