---
title: "CNA vs. Buffa vs. Mitoscore"
author: "Caroline Dravillas"
date: "2025-07-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tmesig)
library(survival)
library(survminer)
library(gridExtra)
library(biomaRt)  
library(cowplot)
```

12 KM curves 

TCC OSTOR
  Filter to IO only 
    - KM NDUFB5 CNA (2 v >2)
    - KM Mitoscore (median)
    - KM Buffa (median)
  Filter to XRT only 
    - KM NDUFB5 CNA (2 v >2)
    - KM Mitoscore (median)
    - KM Buffa (median)
    
ORIEN
  Filter to IO only 
    - KM NDUFB5 CNA (2 v >2)
    - KM Mitoscore (median)
    - KM Buffa (median)
  Filter to XRT only 
    - KM NDUFB5 CNA (2 v >2)
    - KM Mitoscore (median)
    - KM Buffa (median)

# Read in expression data and prep

input should be a df of expression counts where rows are genes, columns are samples,
and the gene symbols are available in a column defined as "Gene".

```{r}
# path <- "/fs/ess/PAS1695/projects/exotho/data/All_Cancer_Gene_expression/"
# files <- list.files(path, pattern = "\\.genes\\.results$", full.names = TRUE)
# 
# combined_df <- map_dfr(files, ~ {
#   df <- read_delim(.x)      # auto-detects header
#   df$sample <- sub("\\.genes\\.results$", "", basename(.x)) # remove suffix
#   df
# })
# 
# raw.expr <- combined_df %>%
#   rename(sample = file_name) %>%
#   select(sample, ) %>%
#   pivot_wider(names_from = sample, values_from = TPM)

raw.expr <- read.delim("/fs/ess/PAS1695/projects/exotho/data/microbiomeRIG/THO_counts.txt")

raw.expr <- raw.expr %>%
  mutate(gene_id = sub("\\..*", "", gene_id))

mart <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
unique_ids <- unique(raw.expr$gene_id)

# Map to gene symbols
gene_map <- getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol"),
  filters = "ensembl_gene_id",
  values = unique_ids,
  mart = mart
)

raw.expr <- raw.expr %>%
  left_join(gene_map, by = c("gene_id" = "ensembl_gene_id"))

raw.expr <- raw.expr %>%
  rename("Gene" = hgnc_symbol) %>%
  dplyr::select(-gene_id)
```

# Calculate Buffa and Mitoscore with tmesig

```{r}
mito <- calculateMitoscore(raw.expr, mito.genes = inputGenes("Mitoscore")) %>%
  mutate(sample = gsub(".genes.results", "", sample))
buffa <- calculateBuffa(raw.expr, buffa.genes = inputGenes("Buffa")) %>%
    mutate(sample = gsub(".genes.results", "", sample))

scores <- mito %>% full_join(buffa)
```

# Read in existing scores 
should be replaced with combined scores calculated above
```{r}
# scores <- read.csv("../data/tho_lung_cancer-mitoBuffa-TPM.csv") 
```

# Read in copy number data 
includes the samples in the two datasets 

```{r}
orien.cna <- readxl::read_xlsx("../data/Whole ORIEN NDUFB5 Copy-Number Values.xlsx") %>%
  mutate(`Sample ID` = SAMPLE_ID) %>%
  #separate(`SAMPLE_ID`, into = c("Pt.id", "sample", "WES.ID"), sep = "_", fill = "right") %>%
  mutate(copy.number = `NDUFB5 Copy Number`) %>%
  dplyr::select(ORIENAvatarKey, `Sample ID`, copy.number) 

tcc.cna <- readxl::read_xlsx("../data/TCC00601-BENE_NDUFB5_copy-number 06022025.xlsx") %>%
  mutate(sample = slid_released,
         copy.number = as.integer(NDUFB5)) %>%
  dplyr::select(sample, copy.number) 
```

# Read in survival data

```{r}
pt.key <- read_delim("../data/tho_lung_cancer.s_clinical_data.tsv", delim = "\t", quote = "", col_names = TRUE) 

clin.clean <- pt.key %>%
  # adjust problem Sample ID 
  mutate(`Sample ID` = ifelse(`Sample ID` == "9Q7N3QFYKC_FT-SA273713D", 
                              "9Q7N3QFYKC_FT-SA273713R_FT-SA273713D",
                              `Sample ID`)) %>%
  mutate(SAMPLE_ID = `Sample ID`) %>%
  separate(`SAMPLE_ID`, into = c("Pt.id", "sample", "WES.ID"), sep = "_", fill = "right") %>%
  mutate(WES.ID = ifelse(`Sequencing Type` == "WES", sample, WES.ID),
         sample = ifelse(`Sequencing Type` == "WES", NA, sample),
         #sample = gsub("-RNA", "", sample),
         vital.status = case_when(PatientVitalStatus == "Dead" ~ 1,
                                  PatientVitalStatus == "Alive" ~ 0),
         AgeAtDeath = ifelse(AgeAtDeath == ">90", 90, AgeAtDeath),
         AgeAtLastContact = ifelse(AgeAtLastContact == ">90", 90, AgeAtLastContact),
         survival.time = case_when(PatientVitalStatus == "Dead" ~ as.double(AgeAtDeath) - as.double(AgeAtDiagnosis),
                                   PatientVitalStatus == "Alive" ~ as.double(AgeAtLastContact) - as.double(AgeAtDiagnosis))) %>%
  rename(ORIENAvatarKey = `Patient ID`) %>%
  filter(!is.na(survival.time)) %>%
  dplyr::select(ORIENAvatarKey, `Sample ID`, sample, WES.ID, vital.status, survival.time, `Sequencing Type`, ImmunoOncologyDrug, RadiationTherapyInd) 
```

# Combine and prep data 

```{r}
# all samples accounted for 
orien.in <- clin.clean %>% 
  left_join(orien.cna) %>%
# include this line if you want to limit it to samples you have CNA for 
  # filter(!is.na(copy.number)) %>%
  left_join(scores) %>%
  mutate(cna.grp = ifelse(copy.number == 2, "Diploid",
                        ifelse(copy.number > 2, "Amplified", NA)),
         mito.grp = ifelse(MitoScore >= median(MitoScore, na.rm = TRUE), "High", "Low"),
         buffa.grp = ifelse(buffa.score >= median(buffa.score, na.rm = TRUE), "High", "Low"))

# 314 samples should have CNA 
tcc.in <- clin.clean %>%
  left_join(tcc.cna) %>%
# include this line if you want to limit it to samples you have CNA for 
  # filter(!is.na(copy.number)) %>%
  left_join(scores) %>%
  filter(sample %in% tcc.cna$sample) %>%
  mutate(cna.grp = ifelse(copy.number == 2, "Diploid",
                        ifelse(copy.number > 2, "Amplified", NA)),
                  mito.grp = ifelse(MitoScore >= median(MitoScore, na.rm = TRUE), "High", "Low"),
         buffa.grp = ifelse(buffa.score >= median(buffa.score, na.rm = TRUE), "High", "Low"))
```

```{r}
km_pairwise_plot <- function(data, group_col_name, filter_col = NULL, filter_value = NULL) {
  subset_data <- data
  subset_data <- subset_data[!is.na(subset_data[[group_col_name]]), ]
  if (!is.null(filter_col)) subset_data <- subset_data[!is.na(subset_data[[filter_col]]), ]
  if (!is.null(filter_col) && !is.null(filter_value)) {
    subset_data <- subset_data[subset_data[[filter_col]] == filter_value, ]
  }

  subset_data$group <- subset_data[[group_col_name]]
  
  fit <- survfit(Surv(survival.time, vital.status) ~ group, data = subset_data)
  
  ggsurvplot(
    fit,
    data = subset_data,
    pval = TRUE,
    risk.table = TRUE,
    legend.title = "Group",
    title = paste(group_col_name,
                  ifelse(!is.null(filter_col), paste("for", filter_col, "== Yes"), "")),
    xlab = "Time",
    xlim = c(0, 20),
    break.time.by = 5,
    palette = "lancet"
  )
}

generate_and_plot_km_curves <- function(data, df_name, group_vars, filter_vars) {
  plots <- list()
  
  for (group_var in group_vars) {
    for (filter_var in filter_vars) {
      plot_name <- paste(df_name, group_var, filter_var, sep = "_")
      p <- km_pairwise_plot(data, group_var, filter_var, "Yes")
      if (!is.null(p)) {
        plots[[plot_name]] <- p
        cat("Plotting:", plot_name, "\n")
        print(p)
      } else {
        message("Skipping plot ", plot_name, " due to no data.")
      }
    }
  }
  
  return(plots)
}

group_vars <- c("cna.grp", "buffa.grp", "mito.grp")
filter_vars <- c("ImmunoOncologyDrug", "RadiationTherapyInd")

# Run for orien.in
orien_plots <- generate_and_plot_km_curves(orien.in, "orien", group_vars, filter_vars)

# Run for tcc.in
tcc_plots <- generate_and_plot_km_curves(tcc.in, "tcc", group_vars, filter_vars)
```


```{r}
combine_survplot <- function(surv_obj) {
  if (is.null(surv_obj)) return(NULL)
  
  plot_grid(
    surv_obj$plot,
    surv_obj$table + theme(plot.margin = margin(0, 0, 0, 0)), 
    ncol = 1,
    align = "v",
    rel_heights = c(2, 0.8)   
  )
}

orien_combined <- lapply(orien_plots, combine_survplot)
tcc_combined   <- lapply(tcc_plots, combine_survplot)
```


```{r}
# Combine orien plots
orien_grid <- grid.arrange(grobs = orien_combined, ncol = 3)

ggsave(
  filename = "../figures/km-curves_orien.png",
  plot = orien_grid,
  width = 20,       # inches
  height = 15,      # inches
  dpi = 300         # high resolution
)

tcc_grid <- grid.arrange(grobs = tcc_combined, ncol = 3)

ggsave(
  filename = "../figures/km-curves_tcc.png",
  plot = tcc_grid,
  width = 20,       # inches
  height = 15,      # inches
  dpi = 300         # high resolution
)
```



 

