---
title: "NDUFB5 CNA Survival curve for diploid vs amplified samples"
author: "C. Dravillas"
date: "2025-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(survival)
library(survminer)
library(gridExtra)
```

groups:
deletion -> cn == <2
diploid -> cn == 2
amplified -> cn >2

Should we include WES and RNA? 
- filter to WES
- well one of the DF's they gave us is only RNAseq and the other has both sample types but only one CN value for both RNAseq and WES?
- Going to use RNAseq ? Since thats what the tcc df is? 
What about multiple samples per patient? 
- filter to primary tumors 1 sample per person
- Don't have this info for all samples
- also this would drop the total to 400 ?
- Maybe I'll do both? 

# Read in copy number and pt key
```{r}
orien <- readxl::read_xlsx("../data/Whole ORIEN NDUFB5 Copy-Number Values.xlsx") %>%
  mutate(`Sample ID` = SAMPLE_ID) %>%
  #separate(`SAMPLE_ID`, into = c("Pt.id", "sample", "WES.ID"), sep = "_", fill = "right") %>%
  mutate(copy.number = `NDUFB5 Copy Number`) %>%
  select(ORIENAvatarKey, `Sample ID`, copy.number)

tcc <- readxl::read_xlsx("../data/TCC00601-BENE_NDUFB5_copy-number 06022025.xlsx") %>%
  mutate(sample = slid_released,
         copy.number = as.integer(NDUFB5)) %>%
  select(sample, copy.number)

pt.key <- read_delim("../data/tho_lung_cancer.s_clinical_data.tsv", delim = "\t", quote = "", col_names = TRUE) 

type <- read.csv("/fs/ess/PAS1695/projects/exorien/data/clinical/20PRJ060OSU_20210712_ClinicalMolLinkage_V4.csv")
```

Clean and prepare data

```{r}
prim.samps <- type %>%
  filter(Disease.Type == "THO - Lung Cancer",
         Primary.Met == "Primary") %>%
  mutate(sample = RNASeq) %>%
  select(sample)

# Unclear if the sample naming is correct 
clin.clean <- pt.key %>%
  mutate(SAMPLE_ID = `Sample ID`) %>%
  separate(`SAMPLE_ID`, into = c("Pt.id", "sample", "WES.ID"), sep = "_", fill = "right") %>%
  # filter to WES only 
  # filter(`Sequencing Type` == "WES" | `Sequencing Type` == "WES, RNASEQ") %>%
  mutate(WES.ID = ifelse(`Sequencing Type` == "WES", sample, WES.ID),
         sample = ifelse(`Sequencing Type` == "WES", NA, sample),
         vital.status = case_when(PatientVitalStatus == "Dead" ~ 1,
                                  PatientVitalStatus == "Alive" ~ 0),
         AgeAtDeath = ifelse(AgeAtDeath == ">90", 90, AgeAtDeath),
         AgeAtLastContact = ifelse(AgeAtLastContact == ">90", 90, AgeAtLastContact),
         survival.time = case_when(PatientVitalStatus == "Dead" ~ as.double(AgeAtDeath) - as.double(AgeAtDiagnosis),
                                   PatientVitalStatus == "Alive" ~ as.double(AgeAtLastContact) - as.double(AgeAtDiagnosis))) %>%
  select(`Patient ID`, `Sample ID`, sample, WES.ID, vital.status, survival.time, `Sequencing Type`) 

orien <- orien %>%
  left_join(clin.clean) %>%
  filter(!is.na(copy.number),
         !is.na(vital.status)) %>%
  select(`Patient ID`, `Sample ID`, sample, WES.ID, `Sequencing Type`, copy.number, vital.status, survival.time)

tcc <- tcc %>%
  left_join(clin.clean) %>%
  filter(!is.na(copy.number),
         !is.na(vital.status)) %>%
  select(`Patient ID`, `Sample ID`, sample, WES.ID, `Sequencing Type`, copy.number, vital.status, survival.time)

# combine and get rid of duplicate rows
surv.in <- rbind(orien, tcc) %>%
  mutate(group = ifelse(copy.number < 2, "Deletion",
                        ifelse(copy.number == 2, "Diploid",
                        ifelse(copy.number > 2, "Amplified", NA))),
         group = ifelse(copy.number > 2 & copy.number < 5, "Mild Amplification",
                        ifelse(copy.number == 2, "Diploid",
                        ifelse(copy.number > 4, "High Amplification", NA)))) %>% 
  distinct()

# can filter to primary samples only but this drops 1255 samples 
surv.in.prim <- surv.in %>% filter(sample %in% prim.samps$sample)

# check duplicate samples n=342
check <- as.data.frame(table(surv.in$`Patient ID`)) %>%
  filter(Freq > 1)

# remove dups
surv.in.filt <- surv.in %>%
  filter(!`Patient ID` %in% check$Var1)
```


##  Plot KM curve 
```{r}
km_pairwise_plot <- function(data, group1, group2) {
  subset_data <- data %>% filter(group %in% c(group1, group2))
  fit <- survfit(Surv(survival.time, vital.status) ~ group, data = subset_data)
  
  # Fit survival model
  fit <- survfit(Surv(survival.time, vital.status) ~ group, data = subset_data)

  # KM plot with axis tweaks
  ggsurvplot(
    fit,
    data = subset_data,
    pval = TRUE,
    risk.table = TRUE,
    legend.title = "Group",
    title = paste(group1, "vs", group2),
    xlab = "Years",
    xlim = c(0, 20),
    palette = "Dark2"
  )
}
```


Plot KM curve - one plot
```{r}
fit <- survfit(Surv(survival.time, vital.status) ~ group, data = surv.in.filt)

# Plot KM curve
ggsurvplot(
  fit,
  data = surv.in,
  pval = TRUE,             
  conf.int = TRUE,         
  risk.table = TRUE,       
  xlab = "Time (years)",  
  ylab = "Survival Probability"
)

pairwise_survdiff(Surv(survival.time, vital.status) ~ group, data = surv.in.filt, p.adjust.method = "BH")
```
3 plots
```{r}
plot1 <- km_pairwise_plot(surv.in.filt, "Amplified", "Diploid")
plot2 <- km_pairwise_plot(surv.in.filt, "Amplified", "Deletion")
plot3 <- km_pairwise_plot(surv.in, "Deletion", "Diploid")

grid.arrange(plot1$plot, plot2$plot, plot3$plot, ncol = 3)
```
Check Mild amplification
```{r}
plot1 <- km_pairwise_plot(surv.in.filt, "Mild Amplification", "Diploid")
plot2 <- km_pairwise_plot(surv.in.filt, "Mild Amplification", "High Amplification")
plot3 <- km_pairwise_plot(surv.in, "High Amplification", "Diploid")

grid.arrange(plot1$plot, plot2$plot, plot3$plot, ncol = 3)
```


<!-- # primary only -->
<!-- # one plot  -->
<!-- ```{r} -->
<!-- fit <- survfit(Surv(survival.time, vital.status) ~ group, data = surv.in.prim) -->

<!-- # Plot KM curve -->
<!-- ggsurvplot( -->
<!--   fit, -->
<!--   data = surv.in.prim, -->
<!--   pval = TRUE,              -->
<!--   conf.int = TRUE,          -->
<!--   risk.table = TRUE,        -->
<!--   xlab = "Time (years)",   -->
<!--   ylab = "Survival Probability" -->
<!-- ) -->

<!-- pairwise_survdiff(Surv(survival.time, vital.status) ~ group, data = surv.in.prim, p.adjust.method = "BH") -->
<!-- ``` -->
<!-- # 3 plots -->
<!-- ```{r} -->
<!-- plot1 <- km_pairwise_plot(surv.in.prim, "Amplified", "Diploid") -->
<!-- plot2 <- km_pairwise_plot(surv.in.prim, "Amplified", "Deletion") -->
<!-- plot3 <- km_pairwise_plot(surv.in.prim, "Deletion", "Diploid") -->

<!-- grid.arrange(plot1$plot, plot2$plot, plot3$plot, ncol = 3) -->
<!-- ``` -->